Network based multifactorial modelling microRNA-target interaction.

### 2019-05-27 article.Rmd
editing second half of introduction

As it is possible for miRNAs to suppress multiple targets, an individual mRNA molecule can also be targeted by multiple miRNAs. In that case, the targeted mRNAs exhibit competitor behavior, that is hypothesized as competing endogenous RNAs (ceRNAs) [@ala_integrated_2013; @cesana_deciphering_2013], against their miRNAs. Briefly, Ala et al. have explained the ceRNA hypothesis as distribution of the other target when a steady-state system that included one miRNA and two target, was perturbed with expression changes at one of the targets [@ala_integrated_2013]. When interaction between miRNAs and their target were dealed in the cell, explanation and prediction of results of perturbation is difficult due to varieties of interactions. Various computational and experimental studies have applied to be understood for ceRNA:miRNA interactions. It has been observed that when abundance of one of the targets of miR-122 was increased, the other target expression also slightly increased as a result of decreasing repressive activity of miR-122 on them [@denzler_assessing_2014]. Bosson et al. have developed a mathematical model for changes on total target pool concentration after grouping to affinity of targets. They have demonstrated that miRNA activity correlated with affinity between miRNA and target [@bosson_endogenous_2014]. Cooperative efficiency of miRNAs as well as competitor behaviors of targets also were studied and it has been demonstrated crucial for regulations at expression of targets [@denzler_impact_2016]. MiRNA:target interactions have modeled as stoichiometric and catalytic mechanisms and Figliuzzi et al. have recommended to handling models in network structures [@figliuzzi_micrornas_2013]. The model that can explain miRNA target interaction through topological features has been applied at bipartite network by Nitzan et al. [@nitzan_interactions_2014]. Robinson and Henderson applied the model that handles miRNA:target direct and indirect interactions via common miRNA of gene and target of miRNA, at bipartite network. It has been demonstrated that whole miRNAs and targets in network can interacted through common miRNAs and genes [@robinson_modelling_2018]. Associated genes that are targets of the same miRNAs have been found with help of correlation of gene expression changes in recent study [@markus_list_sponge_2017]. List et al. have specified that their approach can be useful for ceRNA’s studies and published these approach as a package that can use with R programming.

>>>>

As it is possible for miRNAs to suppress multiple targets, an individual mRNA molecule can also be targeted by multiple miRNAs. In that case, the targeted mRNAs exhibit competitor behavior, that is hypothesized as competing endogenous RNAs (ceRNAs) [@ala_integrated_2013; @cesana_deciphering_2013], against their miRNAs. Briefly, Ala et al. have explained the ceRNA hypothesis as distribution (TOFIX disturbance mı kastedildi acaba?) of the second target when a steady-state system that included one miRNA and two target when perturbed with expression changes at one of the targets [@ala_integrated_2013]. Regarding interaction between miRNAs and their target in a cell, explaining and predicting results of an individual perturbation is difficult due to complexity of interactions. Various computational and experimental studies have tackled the problem of unraveling ceRNA:miRNA interactions. It has been observed that when abundance of one of the targets of miR-122 was increased, the other target expression also slightly increased as a result of decreasing repressive activity of miR-122 on them [@denzler_assessing_2014]. Bosson et al. have developed a mathematical model for changes on total target pool concentration after grouping targets according to affinity (TOFIX orijinal cümle ile aynı manaya geldi mi emin olamadım) and demonstrated that miRNA activity correlated with affinity between miRNA and target [@bosson_endogenous_2014]. Cooperative efficiency of miRNAs as well as competitor behaviors of targets were also studied and it has been demonstrated to be crucial for regulating available mRNA levels (TOFIX expression deyince sanki transcriptional bir etki var gibi anlaşılacak diye rna level deme ihtiyacı hissettim) of targets [@denzler_impact_2016]. MiRNA:target interactions have been modeled as stoichiometric and catalytic mechanisms and Figliuzzi et al. have recommended handling models in network context [@figliuzzi_micrornas_2013]. The model that can explain miRNA target interaction through topological features has been applied at bipartite network by Nitzan et al. [@nitzan_interactions_2014]. Robinson and Henderson applied the model that handles miRNA:target direct and indirect interactions via common miRNA of gene and target of miRNA, at bipartite network. It has been demonstrated that all miRNAs and targets in network can interact each other, bia not necessarily direct interactions, through common miRNAs and genes [@robinson_modelling_2018]. Associated genes that are targets of the same miRNAs have been found with help of correlation of gene expression changes in recent study(TOFIX bu R paketine atıf veriyor, makale atıfı varsa daha iyi olurdu, makalesi yoksa, recent study yerine, recent algoritm deriz) [@markus_list_sponge_2017]. List et al. have specified that their approach can be useful for ceRNA studies and published their approach as an R package.
